---
title: "WEEK 10 HOMEWORK"
author: "Calvin Hsu"
date: "2024-11-16"
output: html_document
---
```{r, warnings=F}
rairuoho <- read.table("rairuoho.txt", header = TRUE, sep = "\t")
library(tidyverse)
```

**subset**
```{r}
enriched_day3 <- rairuoho$length[rairuoho$day == "day3" & rairuoho$treatment == "enriched"]
enriched_day4 <- rairuoho$length[rairuoho$day == "day4" & rairuoho$treatment == "enriched"]
enriched_day8 <- rairuoho$length[rairuoho$day == "day8" & rairuoho$treatment == "enriched"]

water_day3 <- rairuoho$length[rairuoho$day == "day3" & rairuoho$treatment == "water"]
water_day4 <- rairuoho$length[rairuoho$day == "day4" & rairuoho$treatment == "water"]
water_day8 <- rairuoho$length[rairuoho$day == "day8" & rairuoho$treatment == "water"]
```

**make data frame again so plotting will be easier later**
```{r}
df <- data.frame(
  enriched_day3 = enriched_day3,
  enriched_day4 = enriched_day4,
  enriched_day8 = enriched_day8,
  water_day3 = water_day3,
  water_day4 = water_day4,
  water_day8 = water_day8
)
```

**create function (two.sided is default)**
```{r}
pearson_test <- function(x, y, alternative = c("two.sided", "less", "greater")) {
  alternative <- match.arg(alternative)
  n <- length(x)
  r <- cor(x, y)
  df <- n - 2L
  t_value <- sqrt(df) * r/sqrt(1 - r^2)
  PVAL <- switch(alternative, less = pt(t_value, df), 
                 greater = pt(t_value, df, lower.tail = FALSE), 
                 two.sided = 2 * min(pt(t_value, df), pt(t_value, 
                                                         df, lower.tail = FALSE)))
  p_value = as.numeric(PVAL)
  return(list(correlation = r, t = t_value, p_value = p_value))
}
```

**enriched treatment cor test (two sided)**

*day3 & day4*
```{r}
enriched_corr_day34 <- pearson_test(enriched_day3, enriched_day4)
ggplot(data = df, aes(x = enriched_day3 , y = enriched_day4)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  annotate("text", 
           x = 6, 
           y = 15, 
           label = paste0("t = ", round(enriched_corr_day34$t, 2), "\n ", 
                          "p = ", format.pval(enriched_corr_day34$p_value, digits = 3, eps = 1e-10), "\n ",
                          "r = ", round(enriched_corr_day34$correlation, 2)),
           hjust = 0, vjust = 1, size = 5)
```

*day3 & day8*
```{r}
enriched_corr_day38 <- pearson_test(enriched_day3, enriched_day8)
ggplot(data = df, aes(x = enriched_day3 , y = enriched_day8)) + 
  geom_point() +
  annotate("text", 
           x = 6, 
           y = 100, 
           label = paste0("t = ", round(enriched_corr_day38$t, 2), "\n ", 
                          "p = ", round(enriched_corr_day38$p_value, 2), "\n ",
                          "r = ", round(enriched_corr_day38$correlation, 2)),
           hjust = 0, vjust = 1, size = 5)
```

**water treatment cor test (two sided)**

*day3 & day4*
```{r}
water_corr_day34 <- pearson_test(water_day3, water_day4)
ggplot(data = df, aes(x = water_day3 , y = water_day4)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  annotate("text", 
           x = 7, 
           y = 15, 
           label = paste0("t = ", round(water_corr_day34$t, 2), "\n ", 
                          "p = ", format.pval(water_corr_day34$p_value, digits = 3, eps = 1e-10), "\n ",
                          "r = ", round(water_corr_day34$correlation, 2)),
           hjust = 0, vjust = 1, size = 5)  
```

*day3 & day8*
```{r}
water_corr_day38 <- pearson_test(water_day3, water_day8)
ggplot(data = df, aes(x = water_day3 , y = water_day8)) + 
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  annotate("text", 
           x = 7, 
           y = 95, 
           label = paste0("t = ", round(water_corr_day38$t, 2), "\n ", 
                          "p = ", round(water_corr_day38$p_value, 2), "\n ",
                          "r = ", round(water_corr_day38$correlation, 2)),
           hjust = 0, vjust = 1, size = 5)
```

**interpretation**

When the duration of the experiment was not long enough (day3 & day4), both treatments showed positive correlation significantly, r = 0.77 / p = 8.94e-06 for enriched and r = 0.79 / p = 5.2e-06 for water. Probably due to the effect of treatment had not been activated yet. 

When both treatments reached the 8th day, although the correlation was lower compared to the two-day growth (r = 0.46) in water treatment, it was still significant (p = 0.02). But the correlation was very low (r = 0.24) and no significant was found (p = 0.26) in the enriched treatment. Probably due to the plants had absorbed the nutrient so that the change of growth becomes larger. 
