---
title: "WEEK 4 HOMEWORK" 
author: "Calvin Hsu"
date: "2024-10-04"
output: html_document
---

#initiate packages 
```{r}
library(dplyr)
library(tidyr) 
```

#download undone data into r
```{r}
undonereef <- read.table('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt',header=T, sep="\t", dec='.')
```

#STEP1:remove variable "row" and "column"
#STEP2:replace "nutrient" to "enriched" in variable "treatment
#STEP3:combine variable "spatial1" and "spatial2" into "spatial"
#STEP4:combine variable "day3-8" into "day" and add a new variable "length" to get the length of the grass on the given day 
```{r}
donereef <- undonereef %>%
  select(-row , -column)%>%
  mutate(treatment = ifelse(treatment == "nutrient" , "enriched" , "water"))%>%
  unite("spatial" , spatial1 , spatial2 , sep = "_")%>%
  pivot_longer(c(1:6) , names_to = "day", values_to = "length" )
```

#turn column "day" from character into factor 
```{r}
donereef$day <- as.factor(donereef$day)
```

#result
```{r}
donereef
```

#compare to undone data
```{r}
head(undonereef , 10)
```



